Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-07-24 manuscript.Rnw
Removed wordcount from remake, does not work since it is a recursive dependency.
Minor formatting; intro on new page; bit of info in terminology
added a reference

  %texcount manuscript.tex > wordcount.txt

@@

@@

\section*{Introduction}

@@

In this paper we are primarily interested in the distribution of biomass ('partitioning') between leaves and stems, an important component of the residence time of carbon in ecosystems \citep{friend2014}. Previous work based on whole stands \citep{oneilldeangelis1981,enquistniklas2002,reich2014} and individual plants \citep{poorter2012}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). \citet{poorter2012} and \citet{enquistniklas2002} attributed higher LMF in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms, but this begs the question whether differences in LMF are also apparent between deciduous and evergreen functional types within angiosperms. It is also unknown whether PFTs with higher leaf biomass also have higher total leaf area, which is relevant because leaf area drives total light interception and thus productivity. Several classic, oft-cited, studies have assumed, based only on differences in measurements of leaf mass, that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.  
Little is known about global-scale patterns in LMF and LAR in relation to climate. It can be expected that biomass partitioning is correlated with precipitation or mean annual temperature because experiments (cite) and comparisons across sites with contrasting climate \citep{callaway1994,poyatos2007} have shown responsiveness of biomass partitioning to environmental drivers. A recent study \citep{reich2014} demonstrated that stand-scale biomass partitioning was related to mean annual temperature across diverse forest stands. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Notes S1). Again, however, these models do not agree on the degree of plasticity in biomass partitioning, or which climate variables it should respond to.

@@

where $M_i$ is the biomass in some compartment (leaves, stems or roots), $a_i$ the annual allocation of NPP to that compartment, and $\tau_i$ the annual turnover of compartment $i$. It is thus easy to see that partitioning ($M_i/M_T$, where $M_T$ is total biomass) can be different from allocation ($a_i$) because turnover ($\tau_i$) differs between leaf and wood biomass.

@@

\section*{Supporting Information}
Additional supporting information may be found in the online version of this article.\\

>>>>

  %texcount manuscript.tex > wordcount.txt
  %shell("texcount manuscript.tex > wordcount.txt")

@@

%% uncomment this after compiling the ms, then compile again. This is a recursive dependency :(
% Also, requires Perl.
% #' <<>>=
% #' shell("texcount manuscript.tex > wordcount.txt")
% #' nrw <- get_wordcount("wordcount.txt")
% #' @
% #' 
% #' Total number of words: \Sexpr{nrw$Total} \\
% #' Abstract: \Sexpr{nrw$Abstract} \\
% #' Introduction: \Sexpr{nrw$Introduction} \\
% #' Methods and Materials: \Sexpr{nrw$Methods} \\
% #' Results: \Sexpr{nrw$Results} \\
% #' Discussion: \Sexpr{nrw$Discussion} \\
% #' 
% #' Number of figures: 4 (all in colour) \\
% #' Number of tables: 1 \\
% #' Supporting information : 5 Tables, 5 Figures and 1 Notes.

@@

\clearpage
\section*{Introduction}

@@

In this paper we are primarily interested in the distribution of biomass ('partitioning') between leaves and stems, an important component of the residence time of carbon in ecosystems \citep{friend2014}. Previous work based on whole stands \citep{oneilldeangelis1981,enquistniklas2002,reich2014} and individual plants \citep{poorter2012}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). \citet{poorter2012} and \citet{enquistniklas2002} attributed higher LMF in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms, but this begs the question whether differences in LMF are also apparent between deciduous and evergreen functional types within angiosperms. It is also unknown whether PFTs with higher leaf biomass also have higher total leaf area, which is relevant because leaf area drives total light interception and thus productivity. Several classic, oft-cited, studies have assumed, based only on differences in measurements of leaf mass, that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.

Little is known about global-scale patterns in LMF and LAR in relation to climate. It can be expected that biomass partitioning is correlated with precipitation or mean annual temperature because small-scale comparisons across sites with contrasting climate have shown responsiveness of biomass partitioning to environmental drivers \citep{callaway1994,delucia2000,poyatos2007}. A recent study \citep{reich2014} demonstrated that stand-scale biomass partitioning was related to mean annual temperature across diverse forest stands. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Notes S1). Again, however, these models do not agree on the degree of plasticity in biomass partitioning, or which climate variables it should respond to.

@@

where $M_i$ is the biomass in some compartment (leaves, stems or roots), $a_i$ the annual allocation of NPP to that compartment, and $\tau_i$ the annual turnover of compartment $i$. It is thus easy to see that partitioning ($M_i/M_T$, where $M_T$ is total biomass) can be different from allocation ($a_i$) because turnover ($\tau_i$) differs between leaf and wood biomass. Here, we present data on biomass partitioning, which can inform models of allocation only when estimates of turnover are available.

@@

\clearpage
\section*{Supporting Information}
Additional supporting information may be found in the online version of this article.\\
